---
title: "Reproducible Research-CP1"
author: "Shresht Shetty"
date: "7 October 2016"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Reading the data 

I'm going to start off by reading the data.

```{r reading data}
setwd("D://Shresht/My stuff/Science/r/rr/cp1/activityd1")
k <- read.csv("activity.csv")
k$time <- sub("(..)$",":\\1",k$interval)
k$datetime <- as.POSIXct(paste(k$date,k$time), format="%Y-%m-%d %H:%M")
```

##Total number of steps each day

After finding the total number of steps each day,a histogram for the total number of steps is plotted.
We then calculate then mean and median number of steps per day

```{r total steps}
lamp <- aggregate(k$steps,by=list(day=k$date),FUN="sum")## total steps per day
hist(lamp$x,main = "Distribution of total distance covered",xlab = "distance")
cam <- aggregate(k$steps,by=list(day=k$date),FUN="mean")## calculates the mean steps per day
jose <- aggregate(k$steps,by=list(day=k$date),FUN="median")
```

Next,
After that,the time series plot for the 5 minute intervals averaged across all the days are plotted.
```{r time series}
ka <- k[order(k[,3]),]## ordering the data frame based on intervals
kap <- ka[complete.cases(ka$steps),]## removing the rows with missing values
drog <- aggregate(kap$steps,by=list(i=kap$interval),FUN="mean")## calculates the mean steps per interval
##drog$i <- sub("(..)$",":\\1",drog$i)
plot(drog$i,drog$x,type = "l",xlab = "interval",ylab = "mean number of steps")
drog[which.max(drog$x),]

```
We calculate the total number of missing values in the entire set.To then fill in the missing values in the set,we calculate the mean of the interval of which the value belongs to.The average of that interval is filled in for the missing value.
```{r missing values}
kapa <- ka[!complete.cases(ka$steps),] ## data frame with only missing values
length(kapa$steps) ## total number of missing values in the 'steps' vector in the data frame
################# I am going to replace the NA values in each interval by the mean of that interval
library(plyr)
impute.mean <- function(x)replace(x,is.na(x),mean(x,na.rm=TRUE))
ka2 <- ddply(ka,~interval,transform,steps=impute.mean(steps))
ka2 <- ka2[order(ka2$interval),]

```
The code below represents code used for generating a histogram for the total distance covered per day after the missing values are filled in.
```{r new hist}
lamp2 <- aggregate(ka2$steps,by=list(day=ka2$date),FUN="sum")
hist(lamp2$x,main = "Distribution of total distance covered",xlab = "distance")

```
Below is the code for the new mean and median for each interval.
```{r new mean}
cam2 <- aggregate(ka2$steps,by=list(interval=ka2$interval),FUN="mean")
jose2 <- aggregate(ka2$steps,by=list(interval=ka2$interval),FUN="median")

```
Now we are going to crete a time series plot for average distance covered per interval based on whether it's a weekday or weekend.
```{r weekend}
ka2$weekend = chron::is.weekend(ka2$date)
ka21 <- aggregate(ka2$steps~ka2$interval+ka2$weekend,ka2,mean)
names(ka21) <- c("interval","Weekend?","steps")
par(mfrow=c(1,1))
with(ka21,plot(steps~interval,type="n",main="Weekday vs weekend avg"))
with(ka21[ka21$`Weekend?`=="TRUE",],lines(steps~interval),type="l",col="blue")
with(ka21[ka21$`Weekend?`=="FALSE",],lines(steps~interval),type="l",col="16")

```
